---
title: "Spotify Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
    source_code: embed
    logo: Spotify_logo.png 
  runtime: shiny
---

<style>
  .navbar { 
    background-color: #121212 !important; 
    border-color: #121212 !important;
  }
  .navbar-brand {
    color: white !important;
  }

  .navbar-nav > li > a {
    color: white !important;
    transition: color 0.3s ease; /* Smooth transition on hover */
  }

  .navbar-nav > li > a:hover {
    background-color: #1db954 !important; 
  }
  body, .box{
  background-color: #121212 !important;
  }

  .box, .value-box{
    background-color: #1db954 !important;
    color: black !important;
  }

  .navbar-nav > li > a, .navbar-brand {
    color: white !important;
  }
  .navbar .nav > li.active > a, 
  .navbar .nav > li > a:hover {
    background-color: #1db954 !important; /* Spotify Green */
    color: white !important;
  } 
  .navbar .nav > li > a {
    background-color: #212121 !important; /* Dark gray background for other tabs */
    color: white !important;
  }

  .navbar .navbar-nav > li > a {
    border-color: #1db954 !important; /* Spotify Green border */
  }

  .fa-music, .fa-fire, .fa-star, .fa-globe,.fa-headphones, .value-box {
    color: transparant !important;
  }
  .value {
  color: black !important;
  }
  .caption{
  color: #535353;
  }

  .highcharts-title, .highcharts-axis-title, .highcharts-axis-labels, .highcharts-tooltip {
    color: white !important;
  }
  .sidebar{
  background-color: #535353;
  color: #1db954;
  }
  .highcharts text{
  color: white !important;
  }

</style>

```{r setup, include=FALSE}
library(flexdashboard)
library(highcharter)
library(dplyr)
library(readxl)
library(DT)
library(shiny)  

data <- read_excel("Spotify_trendData.xlsx") 

data <- data %>% mutate(
  release_date = as.Date(date),  
  streams = as.numeric(streams),
  `energy_%` = as.numeric(`energy_%`),
  `danceability_%` = as.numeric(`danceability_%`),
  `valence_%` = as.numeric(`valence_%`),
  `acousticness_%` = as.numeric(`acousticness_%`),
  `speechiness_%` = as.numeric(`speechiness_%`)
)


# Define color palette
spotify_green <- "#1db954"
spotify_black <- "#212121"
spotify_dark_gray <- "#535353"
spotify_gray <- "#b3b3b3"
```
Track Analysis
======================================
Row {data-width=250}
----------------------------------------------------------------

### Total Tracks
```{r}
# Total Tracks
total_tracks <- nrow(data)
valueBox(
  value = total_tracks, 
  caption = "Total Tracks", 
  icon = "fa-music", 
  color = spotify_green
)
```

### Most Popular Track
```{r}
# Most Popular Track
popular_track <- data %>% 
  arrange(desc(streams)) %>% 
  slice(1) %>% 
  pull(track_name)
valueBox(
  value = popular_track, 
  caption = "Popular Track", 
  icon = "fa-fire", 
  color = spotify_green
)
```


### Most Popular Artist
```{r}
# Most Popular Artist
popular_artist <- data %>%
  group_by(artist_name) %>%
  summarise(TotalStreams = sum(streams, na.rm = TRUE)) %>%
  arrange(desc(TotalStreams)) %>%
  slice(1) %>%
  pull(artist_name)
valueBox(
  value = popular_artist, 
  caption = "Popular Artist", 
  icon = "fa-star", 
  color = spotify_green
)
```


### Most Popular Artist Across All Playlists
```{r}
# Filter data for tracks released in 2023
data_2023 <- data %>%
  filter(released_year == 2023)

# Find the highest streamed track in 2023
highest_streamed_track_2023 <- data_2023 %>%
  group_by(track_name) %>%
  summarise(TotalStreams = sum(streams, na.rm = TRUE)) %>%
  arrange(desc(TotalStreams)) %>%
  slice(1) %>%
  pull(track_name)

# Display the highest streamed track in 2023 in a value box
valueBox(
  value = highest_streamed_track_2023, 
  caption = "Highest Streamed Track in 2023", 
  icon = "fa-headphones", 
  color = spotify_green
)
```
Row {data-width=500}
-------------------------------------------------------------------------------
### Top Tracks by Stream
```{r}
top_tracks <- data %>%
  arrange(desc(streams)) %>%
  head(10) %>%
  select(track_name, streams)

highchart() %>%
  hc_chart(type = "bar", backgroundColor = "#121212", plotBackgroundColor = "#121212") %>%
  hc_title(text = "Top Tracks by Stream", style = list(color = "#FFFFFF")) %>%
  hc_xAxis(categories = top_tracks$track_name, labels = list(style = list(color = "#FFFFFF"))) %>%
  hc_add_series(name = "Streams", data = top_tracks$streams, color = spotify_green) %>%
  hc_plotOptions(series = list(borderWidth = 0, dataLabels = list(enabled = TRUE, style = list(color = "#FFFFFF")))) %>%
  hc_yAxis(title = list(text = "Streams", style = list(color = "#FFFFFF")))

```
### Top Artists by Total Streams
```{r}
top_artists <- data %>%
  group_by(artist_name) %>%
  summarise(TotalStreams = sum(streams, na.rm = TRUE)) %>%
  arrange(desc(TotalStreams)) %>%
  head(10)

highchart() %>%
  hc_chart(type = "pie", backgroundColor = "#121212", plotBackgroundColor = "#121212") %>%
  hc_title(text = "Top Artists by Total Streams", style = list(color = "#FFFFFF")) %>%
  hc_add_series(
    name = "Total Streams",
    data = list_parse(top_artists %>% select(name = artist_name, y = TotalStreams)),
    colorByPoint = TRUE
  ) %>%
  hc_plotOptions(pie = list(
    dataLabels = list(enabled = TRUE, style = list(color = "#FFFFFF"))
  ))
```

### Stream Distribution by Month
```{r}
monthly_streams <- data %>%
  group_by(released_month) %>%
  summarise(TotalStreams = sum(streams, na.rm = TRUE))

highchart() %>%
  hc_chart(type = "area", backgroundColor = "#121212", plotBackgroundColor = "#121212") %>%
  hc_title(text = "Stream Distribution by Month", style = list(color = "#FFFFFF")) %>%
  hc_xAxis(categories = monthly_streams$released_month, labels = list(style = list(color = "#FFFFFF"))) %>%
  hc_add_series(name = "Total Streams", data = monthly_streams$TotalStreams, color = spotify_green, fillOpacity = 0.3) %>%
  hc_yAxis(title = list(text = "Total Streams", style = list(color = "#FFFFFF")))
```


Advance Analysis
===============================================
Inputs {.sidebar}
----------------------------------------------------------------

```{r}
min_date <- min(data$release_date, na.rm = TRUE)
max_date <- max(data$release_date, na.rm = TRUE)

dateRangeInput("date_range", "Date Range", 
               start = min_date, end = max_date,  # Default to full range
               min = min_date, max = max_date)    # Restrict the range to actual data limits
```

Filtered Data Visualizations {.tabset .tabset-fade}
-----------------------------------------------------

```{r, context="server"}
# Reactive expression to filter data
filtered_data <- reactive({
  req(input$date_range)  # Ensure input is initialized
  data %>%
    filter(
      release_date >= input$date_range[1],
      release_date <= input$date_range[2]
    )
})
```

### Total Streams Over Time
```{r, context="server"}
output$stream_density_calendar <- renderHighchart({
  calendar_data <- filtered_data()

  # Check if filtered data is empty
  if (nrow(calendar_data) == 0) {
    return(highchart() %>% hc_title(text = "No Data Available for Selected Date Range", style = list(color = "red")))
  }

  # Prepare data for calendar plot by counting streams per day
  calendar_data <- calendar_data %>%
    group_by(release_date) %>%
    summarise(TotalStreams = sum(streams, na.rm = TRUE)) %>%
    ungroup()

  highchart() %>%
    hc_chart(type = "scatter", backgroundColor = spotify_black) %>%
    hc_xAxis(type = "datetime", title = list(text = "Date", style = list(color = "#FFFFFF"))) %>%
    hc_yAxis(title = list(text = "Stream Density", style = list(color = "#FFFFFF"))) %>%
    hc_add_series(name = "Daily Streams", data = list_parse2(calendar_data), color = spotify_green, marker = list(radius = 5)) %>%
    hc_tooltip(pointFormat = "{point.x:%Y-%m-%d}: {point.y} streams") %>%
    hc_title(text = "Stream Density by Day (Calendar Plot)", style = list(color = "#FFFFFF"))
})

highchartOutput("stream_density_calendar")
```

### Top 10 Tracks by Streams
```{r, context="server"}
output$top_tracks_wordcloud <- renderHighchart({
  wordcloud_data <- filtered_data()

  # Check if filtered data is empty
  if (nrow(wordcloud_data) == 0) {
    return(highchart() %>% hc_title(text = "No Data Available for Selected Filters", style = list(color = "red")))
  }

  wordcloud_data <- wordcloud_data %>%
    group_by(track_name) %>%
    summarise(TotalStreams = sum(streams, na.rm = TRUE)) %>%
    arrange(desc(TotalStreams)) %>%
    slice_head(n = 10)

  highchart() %>%
    hc_chart(type = "wordcloud", backgroundColor = spotify_black) %>%
    hc_add_series(name = "Streams", data = list_parse2(wordcloud_data), color = spotify_green) %>%
    hc_title(text = "Top 10 Tracks by Streams", style = list(color = "#FFFFFF"))
})

highchartOutput("top_tracks_wordcloud")
```

### Streams by Artist
```{r, context="server"}
output$streams_treemap <- renderHighchart({
  treemap_data <- filtered_data()

  # Check if filtered data is empty
  if (nrow(treemap_data) == 0) {
    return(highchart() %>% hc_title(text = "No Data Available for Selected Date Range", style = list(color = "red")))
  }

  # Group data by artist and calculate total streams
  treemap_data <- treemap_data %>%
    group_by(artist_name) %>%
    summarise(TotalStreams = sum(streams, na.rm = TRUE)) %>%
    arrange(desc(TotalStreams)) %>%
    slice_head(n = 15)  # Limit to top 15 artists for readability

  # Define a color palette for the treemap
  color_palette <- c("#1db954", "#212121", "#121212", "#535353", "#b3b3b3")

  highchart() %>%
    hc_chart(type = "treemap", backgroundColor = spotify_black) %>%
    hc_add_series(
      name = "Streams by Artist",
      data = list_parse(treemap_data %>% transmute(name = artist_name, value = TotalStreams, color = sample(color_palette, n(), replace = TRUE)))
    ) %>%
    hc_title(text = "Streams by Artist", style = list(color = "#FFFFFF"))
})

highchartOutput("streams_treemap")
```

### Distribution of Key Audio Features
```{r, context="server"}
output$audio_features_radar <- renderHighchart({
  radar_data <- filtered_data()

  # Check if filtered data is empty
  if (nrow(radar_data) == 0) {
    return(highchart() %>% hc_title(text = "No Data Available for Selected Filters", style = list(color = "red")))
  }

  radar_data <- radar_data %>%
    summarise(
      Energy = mean(`energy_%`, na.rm = TRUE),
      Danceability = mean(`danceability_%`, na.rm = TRUE),
      Valence = mean(`valence_%`, na.rm = TRUE),
      Acousticness = mean(`acousticness_%`, na.rm = TRUE),
      Speechiness = mean(`speechiness_%`, na.rm = TRUE)
    )

  highchart() %>%
    hc_chart(polar = TRUE, type = "line", backgroundColor = spotify_black) %>%
    hc_xAxis(categories = names(radar_data), tickmarkPlacement = "on", lineWidth = 0, labels = list(style = list(color = "#FFFFFF"))) %>%
    hc_yAxis(gridLineInterpolation = "polygon", lineWidth = 0, min = 0, labels = list(enabled = FALSE)) %>%
    hc_add_series(name = "Audio Features", data = as.numeric(radar_data), color = spotify_green) %>%
    hc_title(text = "Audio Feature Distribution", style = list(color = "#FFFFFF"))
})

highchartOutput("audio_features_radar")
```

Row Data
============================

```{r}
# Render the interactive DataTable
DT::datatable(data, options = list(pageLength = 50, scrollX = TRUE), rownames = FALSE)
```
